### 2021-04-20 Code/Data_Processing.Rmd
Data processing draft

---
title: "Data_Processing"
author: "Katie Krejsa"
date: "4/16/2021"
output: pdf_document
---

## Data Setup

```{r}
# set/check working directory
getwd()

# Import Datasets
IV_tracklogs <- read.csv("./Data/Raw/IV_tracklogs.csv")
IV_wpts <- read.csv("./Data/Raw/IV_wpts.csv")

# Load Packages
library(tidyverse)
library(lubridate)
library(plyr)
library(stringr)
library(tidyr)
library(zoo)

```

## Data Wrangling

```{r}
# Format date as date class in both datasets
IV_tracklogs$date <- as.Date(IV_tracklogs$date, format = "%Y-%m-%d")
IV_wpts$date <- as.Date(IV_wpts$date, format = "%Y-%m-%d")

# Removing the "z" from the wpts dataset
IV_wpts$time1 <- substr(IV_wpts$time, 1, nchar(IV_wpts$time)-1)

# Finding all of the unique dates in the IV_wpts dataset
wpts_dates <- unique(IV_wpts$date)

# Filtering the IV_tracklogs dataset to that it only includes the unique dates found in the IV_wpts dataset
IV_tracklogs_filtered <- IV_tracklogs[IV_tracklogs$date %in% wpts_dates, ]

# Checking that the unique dates in the IV_tacklogs dataset match the unique dates in the IV_wpts dataset
tracklogs_filtered_dates <- unique(IV_tracklogs_filtered$date)
setequal(wpts_dates, tracklogs_filtered_dates)

# Combining the date and time1 in the IV_wpts dataset
IV_wpts$date_time <- as.POSIXct(paste(IV_wpts$date, IV_wpts$time1), format="%Y-%m-%d %H:%M:%S")

# Combining the date and time in the IV_tracklogs_flitered dataset
IV_tracklogs_filtered$date_time <- as.POSIXct(paste(IV_tracklogs_filtered$date, IV_tracklogs_filtered$time), format="%Y-%m-%d %H:%M:%S")

# arranging the column date_time in IV_wpts dataset by ascending order
IV_wpts_ascending <- arrange(IV_wpts, date_time)

# arranging the column date_time in IV_tracklogs_filtered dataset by ascending order
IV_tracklogs_ascending <- arrange(IV_tracklogs_filtered, date_time)

# Creating a column in IV_tracklogs_ascending called "use"
IV_tracklogs_ascending$use <- NA

# Assigning cmt column values to the IV_tracklogs_ascending dataset
for (i in 1:nrow(IV_wpts_ascending)-1){
  IV_tracklogs_ascending$use[IV_tracklogs_ascending$date_time >= IV_wpts_ascending$date_time[i] & IV_tracklogs_ascending$date_time <= IV_wpts_ascending$date_time[i+1]] <- IV_wpts_ascending$cmt[i]
  }

# filtering out all of the rows in the IV_tracklogs_ascending dataset that have "NA" and blank in the "use" column so we are left with only entries that have values in the "use" column
IV_tracklogs_ascending <-
  IV_tracklogs_ascending %>% 
  filter(!is.na(use))

#### Trying to fill in the missing values with previous value - doesn't work with a list/string
#IV_tracklogs_ascending1 <- 
#  IV_tracklogs_ascending %>% 
#  fill(use, use, .direction = 'down')

# Separating the characters and numbers in the "use" column of the IV_tracklogs_ascending dataset to make two new columns
IV_tracklogs_ascending$use_cha <- (str_extract(IV_tracklogs_ascending$use, "[aA-zZ]+"))
IV_tracklogs_ascending$use_num <- as.numeric(str_extract(IV_tracklogs_ascending$use, "[0-9]+"))

# If the "use" column in IV_tracklogs_ascending is blank then the details previously recorded remained the same; Therefore, need to fill in blanks with previously recorded value for use_cha and use_num columns
IV_tracklogs_ascending <- 
  IV_tracklogs_ascending %>% 
  fill(use_cha, use_cha, .direction = 'down') %>% 
  fill(use_num, use_num, .direction = 'down')

# Adding the "name" column from the IV_wpts_ascending dataset to the IV_tracklogs_ascending dataset
IV_tracklogs_ascending$name <- NA

for (i in 1:nrow(IV_wpts_ascending)-1){
  IV_tracklogs_ascending$name[IV_tracklogs_ascending$date_time >= IV_wpts_ascending$date_time[i] & IV_tracklogs_ascending$date_time <= IV_wpts_ascending$date_time[i+1]] <- IV_wpts_ascending$name[i]
  }

```

### 2021-04-20 draft_writing.Rmd
EH_draft_writing

---
title: "draft final Project"
author: "Elise Harrigan"
date: "4/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[knitting](../ENV872_EDA_FinalProject/img/elephant.png)

## ELEPHANT MOVEMENT IN GABON, AFRICA

Elephants are universally some of the most beloved and threatened species in Africa. As a charismatic megafauna, they capture the attention of people as a majestic and powerful animal roaming the African forest and plains. But elephants are also critical in forest ecosystems. As elephant herds travel, they remove trees to create space, consume an incredible amounts of vegetation and use water bodies and dirt piles to cool off. The daily actions of elephants inevitably leave a network of trails in their wake. To be able to better understand forest elephants' daily movements, we will study the forest elephant of Gabon.

Gabon is a country located on the western coast of Africa and covered in dense forest. Due to the forest canopy coverage, collecting data on the elephant trails can be difficult because satellite imagery is not able to penetrate past the top layers of the forest to capture the trail locations on the forest floor. In order to gather more information on elephant movements, data were collected in the field and was analyzed in this study.

## STUDY QUESTIONS

This study focused on understanding the use and frequency of elephant trails. We analyzed the condition of trail as poor, medium or heavy trodden, the width of the trail, and the use denoted by the start, stop or change direction in the trail. We aimed to answer the following questions to gain information on the activity level of elephants and the use of the trails.

1. What is the condition and use of the trail? 
2. What is frequency and number of days traveled on each trail?

By studying and better understanding the movements of elephants, future researchers will be able to compare how elephants react to human disturbance, change in vegetation type and patterns the species make across the landscape.

## DATASETS

The two datasets used are a tracklogs and the waypoints from XX National Park. These data provide the date, time, location, width and type of trail.

__Dataset__          | __Details__
--------------------         |-----------------
IV_tracklogs              | 
IV_wgts      |

### 2021-04-20 Code/Data_Processing.Rmd
Data Processing updates

@@
 library(tidyr)
-library(zoo)
 
@@
 
-#### Trying to fill in the missing values with previous value - doesn't work with a list/string
+#### Attempted code: Trying to fill in the missing values with previous value - doesn't work with a list/string
 #IV_tracklogs_ascending1 <- 
@@
 
+### Attempted code to fill use_num with blanks where use doesnt have a character and fill with previous number when it does. 
+# ifelse(IV_tracklogs_ascending$use != "[0-9]+", IV_tracklogs_ascending$use_num == "", fill(use_num, use_num, .direction = 'down'))
+
 # Adding the "name" column from the IV_wpts_ascending dataset to the IV_tracklogs_ascending dataset
@@
 ```
+
+## Summarizing Data Structure
+
+```{r}
+# Ranges of dates
+range(IV_tracklogs$date)
+range(IV_wpts$date)
+
+# Range and central tendancies of trail widths
+range(IV_tracklogs_ascending$use_num)
+mean(IV_tracklogs_ascending$use_num)
+median(IV_tracklogs_ascending$use_num)
+
+# Range of lattitude and longitude
+range(IV_tracklogs$lat)
+mean(IV_tracklogs$lat)
+median(IV_tracklogs$lat)
+
+range(IV_tracklogs$lon)
+mean(IV_tracklogs$lon)
+median(IV_tracklogs$lon)
+
+# unique values for use_cha
+unique(IV_tracklogs_ascending$use_cha)
+
+```

### 2021-04-20 Code/Exploratory_Analysis.Rmd
commit_EH

---
title: "Exploratory_Analysis"
author: "Elise Harrigan"
date: "4/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up}

getwd()

# Load Packages
library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(leaflet)
library(mapview)
library(raster)
library(sp)
library(rgdal)

#set theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

##load in data - change this to pull from the processed file
tracklogs <- read.csv("../Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("../Data/Processed/IV_wpts_processed.csv")

```
```{r explore the data}
str(tracklogs)
names(tracklogs)
head(tracklogs$lat)
head(tracklogs$lon)

#looking for CRS
st_crs(tracklogs)

```

```{r}
## convert .csv to .shp

 writeOGR(tracklogs, ".", "..Data/Processed/IV_tracklogs_processed.csv", 
           driver = "ESRI Shapefile")

#Read in shapefile into an sf dataframe 
tracklogs_sf <- st_read("../Data/Processed/IV_tracklogs_processed.csv")

#4. Reveal the CRS of the counties features
st_crs(NE_counties_sf)

#5. Plot the data
NE_map <- mapview(NE_counties_sf,
         crs = 4269, 
         grid = FALSE)
NE_map

```

### 2021-04-20 Code/Data_Processing.Rmd
commit_EH
Data Processing updates

@@
 
-#### Attempted code: Trying to fill in the missing values with previous value - doesn't work with a list/string
-#IV_tracklogs_ascending1 <- 
-#  IV_tracklogs_ascending %>% 
-#  fill(use, use, .direction = 'down')
+# If the "use" column in IV_tracklogs_ascending is blank then the details previously recorded remained the same; Therefore, need to fill in blanks with previously recorded value; Need to fill cells of empty strings with NA first before filling
+IV_tracklogs_ascending <-
+  IV_tracklogs_ascending %>% 
+  mutate(use = if_else(use=="", NA_character_,use)) %>% 
+  fill(use, .direction = "down")
 
@@
 IV_tracklogs_ascending$use_num <- as.numeric(str_extract(IV_tracklogs_ascending$use, "[0-9]+"))
-
-# If the "use" column in IV_tracklogs_ascending is blank then the details previously recorded remained the same; Therefore, need to fill in blanks with previously recorded value for use_cha and use_num columns
-IV_tracklogs_ascending <- 
-  IV_tracklogs_ascending %>% 
-  fill(use_cha, use_cha, .direction = 'down') %>% 
-  fill(use_num, use_num, .direction = 'down')
-
-### Attempted code to fill use_num with blanks where use doesnt have a character and fill with previous number when it does. 
-# ifelse(IV_tracklogs_ascending$use != "[0-9]+", IV_tracklogs_ascending$use_num == "", fill(use_num, use_num, .direction = 'down'))
 
@@
 ```
+
+## Saving Processed Datasets
+
+```{r}
+
+write.csv(IV_tracklogs_ascending, row.names = FALSE, file = "./Data/Processed/IV_tracklogs_processed.csv")
+
+write.csv(IV_wpts_ascending, row.names = FALSE, file = "./Data/Processed/IV_wpts_processed.csv")
+
+```

### 2021-04-21 Code/Exploratory_Analysis.Rmd
Exploratory_analysis1

@@
 date: "4/20/2021"
-output: pdf_document
+output: html_document
+geometry: margin=2.54cm
+editor_options:
+  chunk_output_type: console
+  markdown:
+    wrap: 72
 ---
@@
 library(rgdal)
+library(cowplot)
+#install.packages("rnaturalearth")
+#install.packages("rnaturalearthdata")
+library(rnaturalearth)
 
@@
   theme(axis.text = element_text(color = "black"), 
-        legend.position = "top")
+        legend.position = "bottom", legend.justification = "right")
 theme_set(mytheme)
@@
 ##load in data - change this to pull from the processed file
-tracklogs <- read.csv("../Data/Processed/IV_tracklogs_processed.csv")
-wpts <- read.csv("../Data/Processed/IV_wpts_processed.csv")
+tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
+wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")
 
 ```
+
 ```{r explore the data}
@@
 head(tracklogs$lon)
+summary(tracklogs)
 
@@
 st_crs(tracklogs)
+#none found but Amelia said WGS 84. Googled UTM projection and is in Zone 32S
+
+tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
+wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
 
@@
 ```{r}
-## convert .csv to .shp
+#subset based on quality of trail
+Poor_trails <- tracklogs %>%
+  filter(use_cha == "P")
 
- writeOGR(tracklogs, ".", "..Data/Processed/IV_tracklogs_processed.csv", 
-           driver = "ESRI Shapefile")
+Med_trail <- tracklogs %>% 
+  filter(use_cha == "M")
 
-#Read in shapefile into an sf dataframe 
-tracklogs_sf <- st_read("../Data/Processed/IV_tracklogs_processed.csv")
-
-#4. Reveal the CRS of the counties features
-st_crs(NE_counties_sf)
-
-#5. Plot the data
-NE_map <- mapview(NE_counties_sf,
-         crs = 4269, 
-         grid = FALSE)
-NE_map
+##W not H??
+Heavy_trail <- tracklogs %>%
+  filter(use_cha == "W")
 
 ```
+
+```{r how many P, M, W trails}
+
+P_num <- ggplot(tracklogs, aes(y = use_cha)) +
+  geom_bar() + mytheme
+
+print(P_num)
+
+```
+__Condition of Trail __      | __Total Amount__
+--------------------         |---------------------------
+Poor (P)                     | 27,218
+Medium (M)                   | 13,666
+Heavy (H)                    | 26
+
+```{r}
+##plot in ggplots with lat, long
+
+#world is from the rnaturalearth package. gives background for country
+world <- ne_countries(scale = "medium", returnclass = "sf")
+
+##ggplot using world map under
+Trail_gg <- ggplot(data = world) +
+     geom_sf() +
+    geom_point(data = tracklogs, aes(x = lon, y = lat, color = use_cha), size = .4)  + 
+    coord_sf(xlim=c(12.20, 12.90), ylim=c(-0.5, .8), expand = TRUE) +
+  ggtitle("Elephant Trail Points") + 
+  xlab("Longitude") + ylab("Latitude") + 
+  scale_x_continuous(breaks=seq(12, 13, by = .5)) +
+ mytheme
+print(Trail_gg)
+
+## sarah's theme code
+##theme(panel.background = element_rect(fill = "white", colour = "black"))
+
+##PLOT IN MAPVIEW for one date
+
+##filter for dates from 02-28-2018 -- 03-10-2018
+subset_trails <- tracklogs %>%
+  filter(date == "2018-02-28")
+
+#CRS=WGS84 = 4326 or UTM zone 32S = 32732
+
+#convert to sf and include crs (I used UTM zone 32S and coord columns)
+subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lat", "lon"))
+
+#transform to UTM
+trails_UTM <- st_transform(subset_sf, crs=32633)
+
+#Plot the data to see the trails - the projection is off by a little 
+Trail_map1 <- mapview(trails_UTM)
+  Trail_map1
+
+```
+
+## ANALYSIS
+```{r plots by width}
+
+name_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
+  geom_boxplot(fill = "darkblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail") + mytheme
+
+print(name_plots)
+
+```
+
+```{r boxplots trail by widths}
+
+##boxplot width of poor trails
+PT_plot <- ggplot(Poor_trails, aes(x=use_cha, y=use_num)) +
+  geom_boxplot(fill = "seagreen") + 
+  ggtitle("Poor Troddened Trail by widths") + 
+  xlab("Amount of use") + 
+  ylab("Width of trail") + ylim(35, 55) + mytheme
+print(PT_plot)
+
+##boxplot width of medium trails
+MT_plot <- ggplot(Med_trail, aes(x=use_cha, y = use_num)) + 
+  geom_boxplot(fill = "seagreen3") +
+  ggtitle("Medium Troddened Trail by widths") + 
+  xlab("Amount of use") + 
+  ylab("Width of trail") + mytheme
+
+print(MT_plot)
+
+##boxplot width of heavy trails
+HT_plot <- ggplot(Heavy_trail, aes(x=use_cha, y = use_num)) + 
+  geom_boxplot(fill = "seagreen4") +
+  ggtitle("Heavy Troddened Trail by widths") + 
+  xlab("Amount of use") + 
+  ylab("Width of trail") + mytheme
+
+print(HT_plot)
+
+##stack the boxplots
+tstack<- plot_grid(
+  PT_plot,
+  MT_plot,
+  HT_plot,
+align = 'vh', 
+hjust = -1, 
+ncol = 1)
+
+tstack
+
+##boxplot frequency - number of days a trail was used
+
+ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
+  geom_bar() + 
+  scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
+          guide = guide_axis(angle = 45)) + 
+  ggtitle("Amount of Trails by day") +
+  mytheme
+
+```
+
+```{r model - Analysis}
+#is the width of the trail significant on the frequency of use?
+
+#do any of the variables (lat, lon, use_num) explain use_cha
+mod_1 <- lm(data = tracklogs, factor(use_cha) ~ lat + lon + use_num)
+  summary(mod1)
+
+```

### 2021-04-21 draft_writing.Rmd
Added info about datasets to Draft_writing file, small edits to Data_processing file and readme file

@@
 title: "draft final Project"
-author: "Elise Harrigan"
+author: "Elise Harrigan and Katie Krejsa"
 date: "4/19/2021"
@@
 The daily actions of elephants inevitably leave a network of trails in their wake.
-To be able to better understand forest elephants' daily movements, we will study the forest elephant of Gabon.
+To be able to better understand forest elephants' daily movements, we will study the trails used by the forest elephant of Gabon.
 
@@
 
-By studying and better understanding the movements of elephants, future researchers will be able to compare how elephants react to human disturbance, change in vegetation type and patterns the species make across the landscape.
+By studying and better understanding the movements and patterns of elephants, future researchers will be able to compare how elephants react to human disturbance, how they impact the vegetation, and their patterns across the landscape.
 
@@
 
-The two datasets used are a tracklogs and the waypoints from XX National Park. These data provide the date, time, location, width and type of trail.
+The data used for this project were collected in and around Ivindo National Park, Gabon by a team of Duke University researchers, local field guides, and forest peoples.
+Two datasets were provided to us by Dr Amelia Meier of Duke University.
+Data for waypoints and tracklogs were collected in the field with GPS units as the team searched for and walked along forest elephant trails.
+The data are both spatial, containing geographic coordinates, and temporal, containing date and time components.
 
-__Dataset__          | __Details__
---------------------         |-----------------
-IV_tracklogs              | 
-IV_wgts      |
+The tracklogs dataset includes GPS data where the field teams were walking.
+Data contained in this dataset includes the site name (Ivindo), date, time, latitude, and longitude, where everything is recorded in decimal degrees WGS84.
+The dates of the tracklogs range from February 2018 to May 2018.
+
+The waypoints dataset includes the waypoints taken by field teams when they started walking on an elephant trail or if the elephant trail changed characteristics.
+The data contained in this dataset includes date, time, latitude, longitude, trail characteristics (cmt), data on where the elephant trails started and stopped and when the field team got on or of a trail (name), and the symbol used in the GPS (sym).
+More specifically, the "name" field includes "start" or "stop" if the team got on or off a trail, "trstart" or "trstop" if the trail itself started or disappeared, "change" if a characteristic of the trail change, and "jct" if they came to a trail crossing.
+In the cmt field, there are two sets of information: a letter (P = poor, M = medium, H = heavy) representing how heavily trodden the path is and a number representing the width of the trial in cm.
+If the cmt section is blank then the details previously recorded remained the same.
+Data contained in this dataset was also recorded in decimal degrees WGS84 and the dates of data collection also range from February 2018 to May 2018.
+
+__Dataset__  | __Variables__| __Range or Unique Values__| __Central Tendancies__
+-------------|--------------|---------------------------|------------------------
+IV_tracklogs |site          |                           |                                                              -       |date          |2018-02-05 - 2018-05-25    |
+     -       |time          |                           |
+     -       |latitude      |-0.3429768 - 0.6005945     |mean = 0.07083619; median = -0.1230693
+     -       |longitude     |12.43589 - 12.81332        |mean = 12.60479; median = 12.60231
+IV_wgts      |date          |2018-02-06 - 2018-05-25    |
+     -       |time          |                           |
+     -       |latitude      |-0.335521 - 0.599461       |mean = 0.2283088; median = 0.486579
+     -       |longitude     |12.43817 - 12.80200        |mean = 12.66357; median = 12.72618
+     -       |name          |                           |
+     -       |sym           |                           |
+     -       |cmt (letter)  | "P" "M" "W"               |
+     -       |cmt (number)  | 35 - 75                   |mean = 50; median = 47.5

### 2021-04-22 draft_writing.Rmd
Draft Writing edits

@@
 
-This study focused on understanding the use and frequency of elephant trails.
+This study focused on understanding the use of elephant trails and the frequency of travel on these trails.
 We analyzed the condition of trail as poor, medium or heavy trodden, the width of the trail, and the use denoted by the start, stop or change direction in the trail.
@@
 More specifically, the "name" field includes "start" or "stop" if the team got on or off a trail, "trstart" or "trstop" if the trail itself started or disappeared, "change" if a characteristic of the trail change, and "jct" if they came to a trail crossing.
-In the cmt field, there are two sets of information: a letter (P = poor, M = medium, H = heavy) representing how heavily trodden the path is and a number representing the width of the trial in cm.
+In the cmt field, there are two sets of information: a letter (P = poor, M = medium, H = heavy) representing how heavily trodden the path is and a number representing the width of the trail in cm.
 If the cmt section is blank then the details previously recorded remained the same.

### 2021-04-22 draft_writing.Rmd
Updates to the Dataset and Data Wrangling sections

@@
      -       |cmt (number)  | 35 - 75                   |mean = 50; median = 47.5
+
+__Data Wrangling__
+
+The overall goal in wrangling our data was to combine the two files so that all the points on the tracklogs are assigned trail characteristics (on trail, off trail, size and use of trails).
+Doing this would then allow us to quantify the types of trails that elephants use in which contexts.
+
+The tracklogs dataset included a total of 76,098 observations of 6 variables (X, site, date, time, lat, and lon), and the waypoints dataset included a total of 452 observations of 8 variables (X, date, time, lat, lon, name, sym, and cmt).
+To assign trail characteristics (cmt and name) from the waypoints dataset to the tracklogs dataset, we could not do a simple join because the dates and times did not match up exactly.
+So instead, we first identified all of the unique dates in the waypoints dataset.
+We then filtered the tracklogs dataset so that it only included the unique dates found in the waypoints dataset.
+This reduced the number of tracklog observations to 42,179.
+
+Because there are many observations on each day, we next combined the date and time columns to create a combined date_time column in each of the two datasets and arranged both datasets by ascending order of date_time.
+As stated earlier, the two datasets do not have exact matching date_time entires, so to assign cmt column values to the tracklogs dataset based on date_time, we transferred cmt values to the tracklogs dataset if the tracklogs date_time was within a window of the date_time column of the waypoints dataset using a for loop.
+We then filtered out all of the "NA" values of the resulting tracklogs dataset.
+We were told that if the "cmt" column is blank then the details previously recorded remained the same; Therefore, we filled in blanks with previously recorded values by first filling cells of empty strings with "NA" and then using the fill() command.
+Because the cmt column holds two pieces of information (a letter representing how heavily trodden the path is and a number representing the width of the trail), we separated these into two separate columns using the str_extract() command.
+Last, we assigned the "name" column values to the tracklogs dataset in the same way as we assigned the cmt values, using a for loop.

### 2021-04-22 Code/Exploratory_Analysis.Rmd
commit exploratory analysis 4/22

@@
 library(cowplot)
+#install.packages("PMCMRplus")
+library(PMCMRplus)
 #install.packages("rnaturalearth")
@@
 Medium (M)                   | 13,666
-Heavy (H)                    | 26
+Heavy (W)                    | 26
 
@@
 
-##filter for dates from 02-28-2018 -- 03-10-2018
+##filter for dates from 04-15-2018
 subset_trails <- tracklogs %>%
-  filter(date == "2018-02-28")
+  filter(date == "2018-04-15")
 
-#CRS=WGS84 = 4326 or UTM zone 32S = 32732
+##subset for each type of trail for 04-15-2018
+subset_poor <- Poor_trails %>%
+  filter(date == "2018-04-15")
 
-#convert to sf and include crs (I used UTM zone 32S and coord columns)
-subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lat", "lon"))
+subset_med <- Med_trail %>%
+  filter(date == "2018-04-15")
 
-#transform to UTM
-trails_UTM <- st_transform(subset_sf, crs=32633)
+subset_heavy <- Heavy_trail %>%
+  filter(date == "2018-04-15")
 
-#Plot the data to see the trails - the projection is off by a little 
-Trail_map1 <- mapview(trails_UTM)
+#CRS=WGS84 4326
+
+#convert to sf and include crs
+subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lon", "lat"))
+
+subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
+subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
+subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))
+
+#Plot the data to see the trails for 2-28-18 
+Trail_map1 <- mapview(subset_sf, color = use_cha)
   Trail_map1
 
+#Plot so each type of trail is a different color on 4- 15-18 so heavy is included 
+Trail_map2 <- 
+mapview(subset_poor_sf, color = "darkseagreen3", col.regions = "darkseagreen4") +
+  mapview(subset_med_sf, color = "coral2", col.regions = "coral3") +
+    mapview(subset_heavy_sf, color = "lightblue", col.regions = "darkblue")
+Trail_map2
 ```
@@
 ```{r plots by width}
+##Use this plot instead of stacked
 
-name_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
-  geom_boxplot(fill = "darkblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail") + mytheme
+width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
+  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme
 
-print(name_plots)
+print(width_plots)
+
+boxplot(use_num ~ use_cha, data = tracklogs)
+boxplot(count ~ use_cha, data = tracklogs)
 
@@
 ```{r boxplots trail by widths}
-
+## NOT USING THESE STACKED PLOTS
 ##boxplot width of poor trails
@@
   xlab("Amount of use") + 
-  ylab("Width of trail") + ylim(35, 55) + mytheme
+  ylab("Width of trail (cm)") + ylim(35, 55) + mytheme
 print(PT_plot)
@@
 tstack
+```
 
-##boxplot frequency - number of days a trail was used
+```{r boxplot # of days}
+##boxplot count - number of days a trail was used
 
-ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
+freq_plot <- ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
   geom_bar() + 
+  scale_fill_manual(values = c("darkseagreen3", "deepskyblue3", "deeppink")) +
   scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
           guide = guide_axis(angle = 45)) + 
-  ggtitle("Amount of Trails by day") +
+  (theme(legend.position = "right", 
+        legend.text = element_text(size = 12), legend.title = element_text(size = 12))) +
+  labs( y = "Count", 
+        x = "Date", 
+        title = "Number of Trail Points per Day", 
+        fill = "Use of Trail") +
   mytheme
+
+print(freq_plot)
 
@@
 ```{r model - Analysis}
-#is the width of the trail significant on the frequency of use?
 
-#do any of the variables (lat, lon, use_num) explain use_cha
-mod_1 <- lm(data = tracklogs, factor(use_cha) ~ lat + lon + use_num)
-  summary(mod1)
+##which trails have a significantly different width from one another 
+mod3 <- lm(data = tracklogs, use_num ~ use_cha)
+  summary(mod3)
+  TukeyHSD(aov(mod3))
+  plot(TukeyHSD(aov(mod3)))
+
+  plot.design(use_num ~ use_cha, data = tracklogs)
+
+  #results W-M, W-P= sig, P-M=not sig
 

### 2021-04-23 draft_writing.Rmd
commit 4/23

@@
 output: pdf_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 Last, we assigned the "name" column values to the tracklogs dataset in the same way as we assigned the cmt values, using a for loop.
+
+## EXPLORATORY ANALYSIS
+In order to better understand the elephant trail data, we used exploration methods to examine these data and their relationship to one another.
+First, we wanted to see the total amount of trails for each use type (P, M, W) to see which type of trail was most present.
+By summarising each use of trail, we were able to see the Poor trails were the most commonly described trail.
+
+__Use of Trail __            | __Total Amount__
+--------------------         |---------------------------
+Poor (P)                     | 27,218
+Medium (M)                   | 13,666
+Heavy (W)                    | 26
+
+To better visualize these data, we used spatial mapping through mapview to get an idea of where the trails are located and the direction the elephants are traveling.
+Given the size of the dataset, we created a subset of these data to just look at one date where all three trail types were present.
+On April 15, 2018, all three trails were documented.
+
+This map shows how the elephants traveled with each dot representing a point captured.
+It was interesting to see a congregation of points and a change from poorly trodden to medium and heavy trodden trails.
+Further investigation could occur using remote sensing techniques such as lidar, would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.
+
+```{r}
+##Insert code for map
+
+```
+
+## ANALYSIS
+
+A research question for this study was to understand the condition and use of each trail and its level of significance. To answer this we analyzed the widths of each type of trail.
+
+```{r plots by width}
+##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.
+It is evident the poor and medium trails are similar in width and the heavy trails are much wider.
+
+width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
+  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme
+
+print(width_plots)
+
+```
+
+To test our assumptions about the significance of the width of trail to the type, we performed a statistical analysis using a linear regression.
+We ran a post-hoc Tukey test and found that, the heavily trodden trail is significantly wider than the poor and medium trodden trails, but the widths of the poor and medium trodden trails are not significantly different.
+
+```{r}
+mod3 <- lm(data = tracklogs, use_num ~ use_cha)
+  summary(mod3)
+  TukeyHSD(aov(mod3))
+  plot(TukeyHSD(aov(mod3)))
+```
+
+The second question we aimed to answer was to see the number of trail points occurring for each day. To do this, we examined the relationship between the days and the number of trails for each day.
+
+```{r}
+
+##for caption: The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.
+
+freq_plot <- ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
+  geom_bar() + 
+  scale_fill_manual(values = c("darkseagreen3", "deepskyblue3", "deeppink")) +
+  scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
+          guide = guide_axis(angle = 45)) + 
+  (theme(legend.position = "right", 
+        legend.text = element_text(size = 12), legend.title = element_text(size = 12))) +
+  labs( y = "Count", 
+        x = "Date", 
+        title = "Number of Trail Points per Day", 
+        fill = "Use of Trail") +
+  mytheme
+
+print(freq_plot)
+
+```

### 2021-04-26 Code/Exploratory_Analysis.Rmd
commit 4/23
writing and exploratory: 4-26

@@
 summary(tracklogs)
-
+dim(tracklogs)
 #looking for CRS
@@
 print(Trail_gg)
-
-## sarah's theme code
-##theme(panel.background = element_rect(fill = "white", colour = "black"))
 
@@
 
-  plot.design(use_num ~ use_cha, data = tracklogs)
-
   #results W-M, W-P= sig, P-M=not sig

### 2021-04-26 draft_writing.Rmd
Added summary/conclusion description
writing and exploratory: 4-26

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(plyr)
+library(stringr)
+library(tidyr)
+library(lubridate)
+library(tidyverse)
+library(agricolae) 
+library(dplyr)
+library(viridis)
+library(RColorBrewer)
+library(colormap)
+library(sf)
+library(leaflet)
+library(mapview)
+library(raster)
+library(sp)
+library(rgdal)
+library(cowplot)
+#install.packages("PMCMRplus")
+library(PMCMRplus)
+#install.packages("rnaturalearth")
+#install.packages("rnaturalearthdata")
+library(rnaturalearth)
+
+#set theme
+mytheme <- theme_light(base_size = 14) +
+  theme(axis.text = element_text(color = "black"), 
+        legend.position = "bottom", legend.justification = "right")
+theme_set(mytheme)
+
+##load in data - change this to pull from the processed file
+tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
+wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")
+
+#change to data 
+tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
+wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
 ```
@@
 In order to better understand the elephant trail data, we used exploration methods to examine these data and their relationship to one another.
-First, we wanted to see the total amount of trails for each use type (P, M, W) to see which type of trail was most present.
-By summarising each use of trail, we were able to see the Poor trails were the most commonly described trail.
+First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most present.
+By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.
 
@@
 
-To better visualize these data, we used spatial mapping through mapview to get an idea of where the trails are located and the direction the elephants are traveling.
-Given the size of the dataset, we created a subset of these data to just look at one date where all three trail types were present.
-On April 15, 2018, all three trails were documented.
+Visualizing these trails was helpful to explore these data spatially.
+To do this, we used mapview to get an idea of where the trails are located and the type of condition the elephants were traveling on.
+Given the size of the dataset, we created a subset to just look at one date where all three trail types were present.
+On April 15, 2018, all three trails were documented which made this a good date to further examine.
 
-This map shows how the elephants traveled with each dot representing a point captured.
-It was interesting to see a congregation of points and a change from poorly trodden to medium and heavy trodden trails.
-Further investigation could occur using remote sensing techniques such as lidar, would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.
+The map shows how the elephants traveled with each point representing a location captured.
+Here you can also see the majority of the trails are considered poorly trodden until the trails begin to congregate, and we see an increase in the number of medium and heavily trodden trails.
+Further investigation could occur using remote sensing techniques such as lidar, which would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.
 
@@
 ##Insert code for map
+##subset for each type of trail for 04-15-2018
+Poor_trails <- tracklogs %>%
+  filter(use_cha == "P")
 
+Med_trail <- tracklogs %>% 
+  filter(use_cha == "M")
+
+Heavy_trail <- tracklogs %>%
+  filter(use_cha == "W")
+
+#filter by date
+subset_poor <- Poor_trails %>%
+  filter(date == "2018-04-15")
+
+subset_med <- Med_trail %>%
+  filter(date == "2018-04-15")
+
+subset_heavy <- Heavy_trail %>%
+  filter(date == "2018-04-15")
+
+#CRS=WGS84 4326
+
+#convert to sf and include crs
+subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lon", "lat"))
+
+subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
+subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
+subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))
+
+#Plot so each type of trail is a different color on 4- 15-18 so heavy is included 
+Trail_map2 <- 
+mapview(subset_poor_sf, color = "darkseagreen3", col.regions = "darkseagreen4") +
+  mapview(subset_med_sf, color = "coral2", col.regions = "coral3") +
+    mapview(subset_heavy_sf, color = "lightblue", col.regions = "darkblue")
+Trail_map2
 ```
@@
 
-A research question for this study was to understand the condition and use of each trail and its level of significance. To answer this we analyzed the widths of each type of trail.
-
-```{r plots by width}
-##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.
-It is evident the poor and medium trails are similar in width and the heavy trails are much wider.
-
-width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
-  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme
-
-print(width_plots)
-
-```
-
-To test our assumptions about the significance of the width of trail to the type, we performed a statistical analysis using a linear regression.
-We ran a post-hoc Tukey test and found that, the heavily trodden trail is significantly wider than the poor and medium trodden trails, but the widths of the poor and medium trodden trails are not significantly different.
-
-```{r}
-mod3 <- lm(data = tracklogs, use_num ~ use_cha)
-  summary(mod3)
-  TukeyHSD(aov(mod3))
-  plot(TukeyHSD(aov(mod3)))
-```
-
-The second question we aimed to answer was to see the number of trail points occurring for each day. To do this, we examined the relationship between the days and the number of trails for each day.
+A research questions for this study were to gain information on how many trails were occurring each day, the type of condition of the trails and if the width of the trail was statistically significant to the type of use.
+To address the first part of this question, we took a deeper dive into analyzing how many trails were occurring for each day by examining the relationship between the days and the number of trails for each day.
 
@@
 ```
+
+For the second part of our research question, we analyzed the widths and condition of each trail by graphing the relationships.
+It was evident that the heavy trails also had the greatest width and the poor and medium trails had relatively similar widths.
+
+To test our assumptions about the significance of the width of trail to the type, we performed a statistical analysis using a linear regression.
+We ran a post-hoc Tukey test and found that, the heavily trodden trail is significantly wider than the poor and medium trodden trails, but the widths of the poor and medium trodden trails are not significantly different.
+
+```{r plots by width}
+##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.
+It is evident the poor and medium trails are similar in width and the heavy trails are much wider.
+
+width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
+  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme
+
+print(width_plots)
+
+```
+
+```{r}
+mod3 <- lm(data = tracklogs, use_num ~ use_cha)
+  summary(mod3)
+  TukeyHSD(aov(mod3))
+  plot(TukeyHSD(aov(mod3)))
+```

### 2021-04-26 draft_writing.Rmd
draft_recommit

@@
 ## EXPLORATORY ANALYSIS
-In order to better understand the elephant trail data, we used exploration methods to examine these data and their relationship to one another.
+The elephant trail dataset has dimensions of 40,910 observation and 11 variables.
+We looked primarily at the use and the width of each trail and the longitude and latitude.
 First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most present.
@@
 
-Visualizing these trails was helpful to explore these data spatially.
+Visualizing these trails was also helpful to explore these data spatially.
 To do this, we used mapview to get an idea of where the trails are located and the type of condition the elephants were traveling on.

### 2021-04-26 draft_writing.Rmd
Updated Summay and Conclusion Section

@@
 We looked primarily at the use and the width of each trail and the longitude and latitude.
-First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most present.
+First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most common.
 By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.
@@
 To test our assumptions about the significance of the width of trail to the type, we performed a statistical analysis using a linear regression.
-We ran a post-hoc Tukey test and found that, the heavily trodden trail is significantly wider than the poor and medium trodden trails, but the widths of the poor and medium trodden trails are not significantly different.
+We ran a post-hoc Tukey test to determine if the different levels of trail use are statistically different from each other.
 
@@
 ```
+
+### SUMMARY AND CONCLUSIONS
+In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points.
+From our analyses on the widths of each type of trail, we found that the heavily trodden trail is significantly wider than the poor and medium trodden trails and the widths of the poor and medium trodden trails are not significantly different.
+While the heavily trodden trails are significantly wider than the poor and medium trodden trails, it is important to note that the heavily trodden trails have a much smaller number of recorded trail points, which occur in a compact geographic area compared to all of the other recorded points, and only on one day.
+Overall, we can conclude that the majority of recorded trail points are poorly trodden (67%) and medium trodden (33%), and very few are heavily trodden (<0.1%).
+
+The second question we wanted to answer was to see the number of trail points occurring for each day.
+In general, you can see that the level of use increased from poorly trodden to medium trodden over time.
+By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.
+
+Using these elephant trail data for the preliminary analyses we have presented in this project provides an initial exploration into the uses for these data.
+In future studies, these data could be used to see how the trails change due to human disturbance or presence or at how elephants change course or not.
+There are also opportunities with these data to look at the change of vegetation along the heavier trodden trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the opened spaces in the forest.

### 2021-04-27 Code/Project_Template.Rmd
Updates to project template

--- 2021-04-26 draft_writing.Rmd
@@
 ---
-title: "draft final Project"
+output: 
+  pdf_document:
+    keep_tex: yes
+    fig_caption: yes
+    fig_height: 3  
+    fig_width: 4
+    number_sections: yes
+geometry: margin=2.54cm
+title: "Analyzing Forest Elephant Trails in Gabon, Africa"
+subtitle: "https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject"
 author: "Elise Harrigan and Katie Krejsa"
-date: "4/19/2021"
-output: pdf_document
-editor_options: 
-  chunk_output_type: console
+fontsize: 12pt
+mainfont: Times New Roman
+always_allow_html: true
+
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+\newpage
+\tableofcontents 
+\newpage
+\listoffigures 
+\newpage
 
+```{r setup, echo=FALSE, include=FALSE}
+# Set your working directory
+getwd()
+
+# Load your packages
 library(plyr)
@@
 library(sf)
-library(leaflet)
 library(mapview)
@@
 library(cowplot)
-#install.packages("PMCMRplus")
-library(PMCMRplus)
-#install.packages("rnaturalearth")
-#install.packages("rnaturalearthdata")
-library(rnaturalearth)
 
-#set theme
+# Set your ggplot theme
 mytheme <- theme_light(base_size = 14) +
@@
 
-##load in data - change this to pull from the processed file
+# Load your datasets
 tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
@@
 
-#change to data 
+#change the dates
 tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
 wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
+
 ```
 
-[knitting](../ENV872_EDA_FinalProject/img/elephant.png)
+# Rationale and Research Questions
 
-## ELEPHANT MOVEMENT IN GABON, AFRICA
+## Elephant Movement in Gabon, Africa
 
@@
 
-## STUDY QUESTIONS
+## Study Questions
 
@@
 
-1. What is the condition and use of the trail? 
-2. What is frequency and number of days traveled on each trail?
+1. What is frequency and number of days traveled on each trail? 
+2. What is the condition and use of the trail?
 
@@
 
-## DATASETS
+\newpage
+
+# Dataset Information
 
@@
 
-__Data Wrangling__
+## Data Wrangling
 
@@
 
-## EXPLORATORY ANALYSIS
+\newpage
+
+# Exploratory Analysis
+
 The elephant trail dataset has dimensions of 40,910 observation and 11 variables.
@@
 
-```{r}
+```{r, echo=FALSE, fig.cap="Map of all trail use types on April 15, 2018.", fig.align="left", warning=FALSE}
 ##Insert code for map
@@
 #convert to sf and include crs
-subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lon", "lat"))
-
 subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
@@
 
-## ANALYSIS
+\newpage
+
+# Analysis
 
@@
 
-```{r}
+```{r, echo=FALSE, fig.cap="The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.", fig.align="left", warning=FALSE}
 
@@
 
-```{r plots by width}
+```{r plots by width, echo=FALSE, fig.cap="On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.", fig.align="left", warning=FALSE}
 ##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.
@@
 
-```{r}
+```{r, echo=FALSE, fig.cap="Post-hoc Tukey test for trail use.", fig.align="left", warning=FALSE}
 mod3 <- lm(data = tracklogs, use_num ~ use_cha)
-  summary(mod3)
-  TukeyHSD(aov(mod3))
   plot(TukeyHSD(aov(mod3)))
@@
 
-### SUMMARY AND CONCLUSIONS
+\newpage
+
+# Summary and Conclusions
+
+The first question we wanted to answer was to see the number of trail points occurring for each day.
+In general, you can see that the level of use increased from poorly trodden to medium trodden over time.
+By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.
+
 In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points.
@@
 
-The second question we wanted to answer was to see the number of trail points occurring for each day.
-In general, you can see that the level of use increased from poorly trodden to medium trodden over time.
-By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.
-
-Using these elephant trail data for the preliminary analyses we have presented in this project provides an initial exploration into the uses for these data.
-In future studies, these data could be used to see how the trails change due to human disturbance or presence or at how elephants change course or not.
-There are also opportunities with these data to look at the change of vegetation along the heavier trodden trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the opened spaces in the forest.
+The preliminary analyses presented in this project, provide an initial exploration into the uses for these data.
+In future studies, these data could be used to see how the trails change due to human disturbance or presence, or to see how and where elephants do or do not change course.
+There are also opportunities with these data to look at the change of vegetation along the different use trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the forest gaps.

### 2021-04-27 Code/Project_Template.tex
Updates to project template

--- 2021-04-27 Code/Project_Template.Rmd
@@
----
-output: 
-  pdf_document:
-    keep_tex: yes
-    fig_caption: yes
-    fig_height: 3  
-    fig_width: 4
-    number_sections: yes
-geometry: margin=2.54cm
-title: "Analyzing Forest Elephant Trails in Gabon, Africa"
-subtitle: "https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject"
-author: "Elise Harrigan and Katie Krejsa"
-fontsize: 12pt
-mainfont: Times New Roman
-always_allow_html: true
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+    \setmainfont[]{Times New Roman}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={Analyzing Forest Elephant Trails in Gabon, Africa},
+            pdfauthor={Elise Harrigan and Katie Krejsa},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage[margin=2.54cm]{geometry}
+\usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
----
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+
+\title{Analyzing Forest Elephant Trails in Gabon, Africa}
+\providecommand{\subtitle}[1]{}
+\subtitle{\url{https://github.com/kmk85/HarriganKrejsa_ENV872_EDA_FinalProject}}
+\author{Elise Harrigan and Katie Krejsa}
+\date{}
+
+\begin{document}
+\maketitle
 
@@
 
-```{r setup, echo=FALSE, include=FALSE}
-# Set your working directory
-getwd()
+\hypertarget{rationale-and-research-questions}{%
+\section{Rationale and Research
+Questions}\label{rationale-and-research-questions}}
 
-# Load your packages
-library(plyr)
-library(stringr)
-library(tidyr)
-library(lubridate)
-library(tidyverse)
-library(agricolae) 
-library(dplyr)
-library(viridis)
-library(RColorBrewer)
-library(colormap)
-library(sf)
-library(mapview)
-library(raster)
-library(sp)
-library(rgdal)
-library(cowplot)
+\hypertarget{elephant-movement-in-gabon-africa}{%
+\subsection{Elephant Movement in Gabon,
+Africa}\label{elephant-movement-in-gabon-africa}}
 
-# Set your ggplot theme
-mytheme <- theme_light(base_size = 14) +
-  theme(axis.text = element_text(color = "black"), 
-        legend.position = "bottom", legend.justification = "right")
-theme_set(mytheme)
+Elephants are universally some of the most beloved and threatened
+species in Africa. As a charismatic megafauna, they capture the
+attention of people as a majestic and powerful animal roaming the
+African forest and plains. But elephants are also critical in forest
+ecosystems. As elephant herds travel, they remove trees to create space,
+consume an incredible amounts of vegetation and use water bodies and
+dirt piles to cool off. The daily actions of elephants inevitably leave
+a network of trails in their wake. To be able to better understand
+forest elephants' daily movements, we will study the trails used by the
+forest elephant of Gabon.
 
-# Load your datasets
-tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
-wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")
+Gabon is a country located on the western coast of Africa and covered in
+dense forest. Due to the forest canopy coverage, collecting data on the
+elephant trails can be difficult because satellite imagery is not able
+to penetrate past the top layers of the forest to capture the trail
+locations on the forest floor. In order to gather more information on
+elephant movements, data were collected in the field and was analyzed in
+this study.
 
-#change the dates
-tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
-wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
+\hypertarget{study-questions}{%
+\subsection{Study Questions}\label{study-questions}}
 
-```
+This study focused on understanding the use of elephant trails and the
+frequency of travel on these trails. We analyzed the condition of trail
+as poor, medium or heavy trodden, the width of the trail, and the use
+denoted by the start, stop or change direction in the trail. We aimed to
+answer the following questions to gain information on the activity level
+of elephants and the use of the trails.
 
-# Rationale and Research Questions
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  What is frequency and number of days traveled on each trail?
+\item
+  What is the condition and use of the trail?
+\end{enumerate}
 
-## Elephant Movement in Gabon, Africa
-
-Elephants are universally some of the most beloved and threatened species in Africa.
-As a charismatic megafauna, they capture the attention of people as a majestic and powerful animal roaming the African forest and plains.
-But elephants are also critical in forest ecosystems.
-As elephant herds travel, they remove trees to create space, consume an incredible amounts of vegetation and use water bodies and dirt piles to cool off.
-The daily actions of elephants inevitably leave a network of trails in their wake.
-To be able to better understand forest elephants' daily movements, we will study the trails used by the forest elephant of Gabon.
-
-Gabon is a country located on the western coast of Africa and covered in dense forest.
-Due to the forest canopy coverage, collecting data on the elephant trails can be difficult because satellite imagery is not able to penetrate past the top layers of the forest to capture the trail locations on the forest floor.
-In order to gather more information on elephant movements, data were collected in the field and was analyzed in this study.
-
-## Study Questions
-
-This study focused on understanding the use of elephant trails and the frequency of travel on these trails.
-We analyzed the condition of trail as poor, medium or heavy trodden, the width of the trail, and the use denoted by the start, stop or change direction in the trail.
-We aimed to answer the following questions to gain information on the activity level of elephants and the use of the trails.
-
-1. What is frequency and number of days traveled on each trail? 
-2. What is the condition and use of the trail?
-
-By studying and better understanding the movements and patterns of elephants, future researchers will be able to compare how elephants react to human disturbance, how they impact the vegetation, and their patterns across the landscape.
+By studying and better understanding the movements and patterns of
+elephants, future researchers will be able to compare how elephants
+react to human disturbance, how they impact the vegetation, and their
+patterns across the landscape.
 
@@
 
-# Dataset Information
+\hypertarget{dataset-information}{%
+\section{Dataset Information}\label{dataset-information}}
 
-The data used for this project were collected in and around Ivindo National Park, Gabon by a team of Duke University researchers, local field guides, and forest peoples.
-Two datasets were provided to us by Dr Amelia Meier of Duke University.
-Data for waypoints and tracklogs were collected in the field with GPS units as the team searched for and walked along forest elephant trails.
-The data are both spatial, containing geographic coordinates, and temporal, containing date and time components.
+The data used for this project were collected in and around Ivindo
+National Park, Gabon by a team of Duke University researchers, local
+field guides, and forest peoples. Two datasets were provided to us by Dr
+Amelia Meier of Duke University. Data for waypoints and tracklogs were
+collected in the field with GPS units as the team searched for and
+walked along forest elephant trails. The data are both spatial,
+containing geographic coordinates, and temporal, containing date and
+time components.
 
-The tracklogs dataset includes GPS data where the field teams were walking.
-Data contained in this dataset includes the site name (Ivindo), date, time, latitude, and longitude, where everything is recorded in decimal degrees WGS84.
-The dates of the tracklogs range from February 2018 to May 2018.
+The tracklogs dataset includes GPS data where the field teams were
+walking. Data contained in this dataset includes the site name (Ivindo),
+date, time, latitude, and longitude, where everything is recorded in
+decimal degrees WGS84. The dates of the tracklogs range from February
+2018 to May 2018.
 
-The waypoints dataset includes the waypoints taken by field teams when they started walking on an elephant trail or if the elephant trail changed characteristics.
-The data contained in this dataset includes date, time, latitude, longitude, trail characteristics (cmt), data on where the elephant trails started and stopped and when the field team got on or of a trail (name), and the symbol used in the GPS (sym).
-More specifically, the "name" field includes "start" or "stop" if the team got on or off a trail, "trstart" or "trstop" if the trail itself started or disappeared, "change" if a characteristic of the trail change, and "jct" if they came to a trail crossing.
-In the cmt field, there are two sets of information: a letter (P = poor, M = medium, H = heavy) representing how heavily trodden the path is and a number representing the width of the trail in cm.
-If the cmt section is blank then the details previously recorded remained the same.
-Data contained in this dataset was also recorded in decimal degrees WGS84 and the dates of data collection also range from February 2018 to May 2018.
+The waypoints dataset includes the waypoints taken by field teams when
+they started walking on an elephant trail or if the elephant trail
+changed characteristics. The data contained in this dataset includes
+date, time, latitude, longitude, trail characteristics (cmt), data on
+where the elephant trails started and stopped and when the field team
+got on or of a trail (name), and the symbol used in the GPS (sym). More
+specifically, the ``name'' field includes ``start'' or ``stop'' if the
+team got on or off a trail, ``trstart'' or ``trstop'' if the trail
+itself started or disappeared, ``change'' if a characteristic of the
+trail change, and ``jct'' if they came to a trail crossing. In the cmt
+field, there are two sets of information: a letter (P = poor, M =
+medium, H = heavy) representing how heavily trodden the path is and a
+number representing the width of the trail in cm. If the cmt section is
+blank then the details previously recorded remained the same. Data
+contained in this dataset was also recorded in decimal degrees WGS84 and
+the dates of data collection also range from February 2018 to May 2018.
 
-__Dataset__  | __Variables__| __Range or Unique Values__| __Central Tendancies__
--------------|--------------|---------------------------|------------------------
-IV_tracklogs |site          |                           |                                                              -       |date          |2018-02-05 - 2018-05-25    |
-     -       |time          |                           |
-     -       |latitude      |-0.3429768 - 0.6005945     |mean = 0.07083619; median = -0.1230693
-     -       |longitude     |12.43589 - 12.81332        |mean = 12.60479; median = 12.60231
-IV_wgts      |date          |2018-02-06 - 2018-05-25    |
-     -       |time          |                           |
-     -       |latitude      |-0.335521 - 0.599461       |mean = 0.2283088; median = 0.486579
-     -       |longitude     |12.43817 - 12.80200        |mean = 12.66357; median = 12.72618
-     -       |name          |                           |
-     -       |sym           |                           |
-     -       |cmt (letter)  | "P" "M" "W"               |
-     -       |cmt (number)  | 35 - 75                   |mean = 50; median = 47.5
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+\textbf{Dataset} & \textbf{Variables} & \textbf{Range or Unique Values}
+& \textbf{Central Tendancies}\tabularnewline
+\midrule
+\endhead
+IV\_tracklogs & site & & -\tabularnewline
+- & time & &\tabularnewline
+- & latitude & -0.3429768 - 0.6005945 & mean = 0.07083619; median =
+-0.1230693\tabularnewline
+- & longitude & 12.43589 - 12.81332 & mean = 12.60479; median =
+12.60231\tabularnewline
+IV\_wgts & date & 2018-02-06 - 2018-05-25 &\tabularnewline
+- & time & &\tabularnewline
+- & latitude & -0.335521 - 0.599461 & mean = 0.2283088; median =
+0.486579\tabularnewline
+- & longitude & 12.43817 - 12.80200 & mean = 12.66357; median =
+12.72618\tabularnewline
+- & name & &\tabularnewline
+- & sym & &\tabularnewline
+- & cmt (letter) & ``P'' ``M'' ``W'' &\tabularnewline
+- & cmt (number) & 35 - 75 & mean = 50; median = 47.5\tabularnewline
+\bottomrule
+\end{longtable}
 
-## Data Wrangling
+\hypertarget{data-wrangling}{%
+\subsection{Data Wrangling}\label{data-wrangling}}
 
-The overall goal in wrangling our data was to combine the two files so that all the points on the tracklogs are assigned trail characteristics (on trail, off trail, size and use of trails).
-Doing this would then allow us to quantify the types of trails that elephants use in which contexts.
+The overall goal in wrangling our data was to combine the two files so
+that all the points on the tracklogs are assigned trail characteristics
+(on trail, off trail, size and use of trails). Doing this would then
+allow us to quantify the types of trails that elephants use in which
+contexts.
 
-The tracklogs dataset included a total of 76,098 observations of 6 variables (X, site, date, time, lat, and lon), and the waypoints dataset included a total of 452 observations of 8 variables (X, date, time, lat, lon, name, sym, and cmt).
-To assign trail characteristics (cmt and name) from the waypoints dataset to the tracklogs dataset, we could not do a simple join because the dates and times did not match up exactly.
-So instead, we first identified all of the unique dates in the waypoints dataset.
-We then filtered the tracklogs dataset so that it only included the unique dates found in the waypoints dataset.
-This reduced the number of tracklog observations to 42,179.
+The tracklogs dataset included a total of 76,098 observations of 6
+variables (X, site, date, time, lat, and lon), and the waypoints dataset
+included a total of 452 observations of 8 variables (X, date, time, lat,
+lon, name, sym, and cmt). To assign trail characteristics (cmt and name)
+from the waypoints dataset to the tracklogs dataset, we could not do a
+simple join because the dates and times did not match up exactly. So
+instead, we first identified all of the unique dates in the waypoints
+dataset. We then filtered the tracklogs dataset so that it only included
+the unique dates found in the waypoints dataset. This reduced the number
+of tracklog observations to 42,179.
 
-Because there are many observations on each day, we next combined the date and time columns to create a combined date_time column in each of the two datasets and arranged both datasets by ascending order of date_time.
-As stated earlier, the two datasets do not have exact matching date_time entires, so to assign cmt column values to the tracklogs dataset based on date_time, we transferred cmt values to the tracklogs dataset if the tracklogs date_time was within a window of the date_time column of the waypoints dataset using a for loop.
-We then filtered out all of the "NA" values of the resulting tracklogs dataset.
-We were told that if the "cmt" column is blank then the details previously recorded remained the same; Therefore, we filled in blanks with previously recorded values by first filling cells of empty strings with "NA" and then using the fill() command.
-Because the cmt column holds two pieces of information (a letter representing how heavily trodden the path is and a number representing the width of the trail), we separated these into two separate columns using the str_extract() command.
-Last, we assigned the "name" column values to the tracklogs dataset in the same way as we assigned the cmt values, using a for loop.
+Because there are many observations on each day, we next combined the
+date and time columns to create a combined date\_time column in each of
+the two datasets and arranged both datasets by ascending order of
+date\_time. As stated earlier, the two datasets do not have exact
+matching date\_time entires, so to assign cmt column values to the
+tracklogs dataset based on date\_time, we transferred cmt values to the
+tracklogs dataset if the tracklogs date\_time was within a window of the
+date\_time column of the waypoints dataset using a for loop. We then
+filtered out all of the ``NA'' values of the resulting tracklogs
+dataset. We were told that if the ``cmt'' column is blank then the
+details previously recorded remained the same; Therefore, we filled in
+blanks with previously recorded values by first filling cells of empty
+strings with ``NA'' and then using the fill() command. Because the cmt
+column holds two pieces of information (a letter representing how
+heavily trodden the path is and a number representing the width of the
+trail), we separated these into two separate columns using the
+str\_extract() command. Last, we assigned the ``name'' column values to
+the tracklogs dataset in the same way as we assigned the cmt values,
+using a for loop.
 
@@
 
-# Exploratory Analysis
+\hypertarget{exploratory-analysis}{%
+\section{Exploratory Analysis}\label{exploratory-analysis}}
 
-The elephant trail dataset has dimensions of 40,910 observation and 11 variables.
-We looked primarily at the use and the width of each trail and the longitude and latitude.
-First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most common.
-By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.
+The elephant trail dataset has dimensions of 40,910 observation and 11
+variables. We looked primarily at the use and the width of each trail
+and the longitude and latitude. First, we wanted to see the total amount
+of trails for each use type (P, M, W) in order to see which type of
+trail was most common. By summarizing the total amounts of each use of
+trail, we were able to see the Poor trails were the most commonly
+described trail.
 
-__Use of Trail __            | __Total Amount__
---------------------         |---------------------------
-Poor (P)                     | 27,218
-Medium (M)                   | 13,666
-Heavy (W)                    | 26
+\begin{longtable}[]{@{}ll@{}}
+\toprule
+\textbf{Use of Trail } & \textbf{Total Amount}\tabularnewline
+\midrule
+\endhead
+Poor (P) & 27,218\tabularnewline
+Medium (M) & 13,666\tabularnewline
+Heavy (W) & 26\tabularnewline
+\bottomrule
+\end{longtable}
 
-Visualizing these trails was also helpful to explore these data spatially.
-To do this, we used mapview to get an idea of where the trails are located and the type of condition the elephants were traveling on.
-Given the size of the dataset, we created a subset to just look at one date where all three trail types were present.
-On April 15, 2018, all three trails were documented which made this a good date to further examine.
+Visualizing these trails was also helpful to explore these data
+spatially. To do this, we used mapview to get an idea of where the
+trails are located and the type of condition the elephants were
+traveling on. Given the size of the dataset, we created a subset to just
+look at one date where all three trail types were present. On April 15,
+2018, all three trails were documented which made this a good date to
+further examine.
 
-The map shows how the elephants traveled with each point representing a location captured.
-Here you can also see the majority of the trails are considered poorly trodden until the trails begin to congregate, and we see an increase in the number of medium and heavily trodden trails.
-Further investigation could occur using remote sensing techniques such as lidar, which would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.
+The map shows how the elephants traveled with each point representing a
+location captured. Here you can also see the majority of the trails are
+considered poorly trodden until the trails begin to congregate, and we
+see an increase in the number of medium and heavily trodden trails.
+Further investigation could occur using remote sensing techniques such
+as lidar, which would allow us to see if there is a environmental
+feature such as a water body, bringing the elephants to this area.
 
-```{r, echo=FALSE, fig.cap="Map of all trail use types on April 15, 2018.", fig.align="left", warning=FALSE}
-##Insert code for map
-##subset for each type of trail for 04-15-2018
-Poor_trails <- tracklogs %>%
-  filter(use_cha == "P")
-
-Med_trail <- tracklogs %>% 
-  filter(use_cha == "M")
-
-Heavy_trail <- tracklogs %>%
-  filter(use_cha == "W")
-
-#filter by date
-subset_poor <- Poor_trails %>%
-  filter(date == "2018-04-15")
-
-subset_med <- Med_trail %>%
-  filter(date == "2018-04-15")
-
-subset_heavy <- Heavy_trail %>%
-  filter(date == "2018-04-15")
-
-#CRS=WGS84 4326
-
-#convert to sf and include crs
-subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
-subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
-subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))
-
-#Plot so each type of trail is a different color on 4- 15-18 so heavy is included 
-Trail_map2 <- 
-mapview(subset_poor_sf, color = "darkseagreen3", col.regions = "darkseagreen4") +
-  mapview(subset_med_sf, color = "coral2", col.regions = "coral3") +
-    mapview(subset_heavy_sf, color = "lightblue", col.regions = "darkblue")
-Trail_map2
-```
+\begin{verbatim}
+## PhantomJS not found. You can install it with webshot::install_phantomjs(). If it is installed, please make sure the phantomjs executable can be found via the PATH variable.
+\end{verbatim}
 
@@
 
-# Analysis
+\hypertarget{analysis}{%
+\section{Analysis}\label{analysis}}
 
-A research questions for this study were to gain information on how many trails were occurring each day, the type of condition of the trails and if the width of the trail was statistically significant to the type of use.
-To address the first part of this question, we took a deeper dive into analyzing how many trails were occurring for each day by examining the relationship between the days and the number of trails for each day.
+A research questions for this study were to gain information on how many
+trails were occurring each day, the type of condition of the trails and
+if the width of the trail was statistically significant to the type of
+use. To address the first part of this question, we took a deeper dive
+into analyzing how many trails were occurring for each day by examining
+the relationship between the days and the number of trails for each day.
 
-```{r, echo=FALSE, fig.cap="The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.", fig.align="left", warning=FALSE}
+\begin{figure}
 
-##for caption: The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.
+\includegraphics{Project_Template_files/figure-latex/unnamed-chunk-2-1} \hfill{}
 
-freq_plot <- ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
-  geom_bar() + 
-  scale_fill_manual(values = c("darkseagreen3", "deepskyblue3", "deeppink")) +
-  scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
-          guide = guide_axis(angle = 45)) + 
-  (theme(legend.position = "right", 
-        legend.text = element_text(size = 12), legend.title = element_text(size = 12))) +
-  labs( y = "Count", 
-        x = "Date", 
-        title = "Number of Trail Points per Day", 
-        fill = "Use of Trail") +
-  mytheme
+\caption{The poor trodden trails are more abundant earlier in the year and the medium trodden trails take over in late April-May timeframe.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
-print(freq_plot)
+For the second part of our research question, we analyzed the widths and
+condition of each trail by graphing the relationships. It was evident
+that the heavy trails also had the greatest width and the poor and
+medium trails had relatively similar widths.
 
-```
+To test our assumptions about the significance of the width of trail to
+the type, we performed a statistical analysis using a linear regression.
+We ran a post-hoc Tukey test to determine if the different levels of
+trail use are statistically different from each other.
 
-For the second part of our research question, we analyzed the widths and condition of each trail by graphing the relationships.
-It was evident that the heavy trails also had the greatest width and the poor and medium trails had relatively similar widths.
+\begin{figure}
 
-To test our assumptions about the significance of the width of trail to the type, we performed a statistical analysis using a linear regression.
-We ran a post-hoc Tukey test to determine if the different levels of trail use are statistically different from each other.
+\includegraphics{Project_Template_files/figure-latex/plots by width-1} \hfill{}
 
-```{r plots by width, echo=FALSE, fig.cap="On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.", fig.align="left", warning=FALSE}
-##for caption: On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.
-It is evident the poor and medium trails are similar in width and the heavy trails are much wider.
+\caption{On the x axis, the three letters represent how heavily trodden the path is:  (P = poor, M = medium, W = heavy) and the width of the trail on the y-axis in centimeters.}\label{fig:plots by width}
+\end{figure}
 
-width_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
-  geom_boxplot(fill = "lightblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail (cm)") + mytheme
+\begin{figure}
 
-print(width_plots)
+\includegraphics{Project_Template_files/figure-latex/unnamed-chunk-3-1} \hfill{}
 
-```
-
-```{r, echo=FALSE, fig.cap="Post-hoc Tukey test for trail use.", fig.align="left", warning=FALSE}
-mod3 <- lm(data = tracklogs, use_num ~ use_cha)
-  plot(TukeyHSD(aov(mod3)))
-```
+\caption{Post-hoc Tukey test for trail use.}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 
-# Summary and Conclusions
+\hypertarget{summary-and-conclusions}{%
+\section{Summary and Conclusions}\label{summary-and-conclusions}}
 
-The first question we wanted to answer was to see the number of trail points occurring for each day.
-In general, you can see that the level of use increased from poorly trodden to medium trodden over time.
-By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.
+The first question we wanted to answer was to see the number of trail
+points occurring for each day. In general, you can see that the level of
+use increased from poorly trodden to medium trodden over time. By
+looking at the graph, the poor trodden trails are more abundant earlier
+in the year and the medium trodden trails become more abundant in April
+and May.
 
-In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points.
-From our analyses on the widths of each type of trail, we found that the heavily trodden trail is significantly wider than the poor and medium trodden trails and the widths of the poor and medium trodden trails are not significantly different.
-While the heavily trodden trails are significantly wider than the poor and medium trodden trails, it is important to note that the heavily trodden trails have a much smaller number of recorded trail points, which occur in a compact geographic area compared to all of the other recorded points, and only on one day.
-Overall, we can conclude that the majority of recorded trail points are poorly trodden (67%) and medium trodden (33%), and very few are heavily trodden (<0.1%).
+In our final dataset, which included only the data that were recorded on
+common dates between our two initial datasets, there are 27,218 poor
+trodden trail points, 13,666 medium trodden trail points, and 26 heavily
+trodden trail points. From our analyses on the widths of each type of
+trail, we found that the heavily trodden trail is significantly wider
+than the poor and medium trodden trails and the widths of the poor and
+medium trodden trails are not significantly different. While the heavily
+trodden trails are significantly wider than the poor and medium trodden
+trails, it is important to note that the heavily trodden trails have a
+much smaller number of recorded trail points, which occur in a compact
+geographic area compared to all of the other recorded points, and only
+on one day. Overall, we can conclude that the majority of recorded trail
+points are poorly trodden (67\%) and medium trodden (33\%), and very few
+are heavily trodden (\textless{}0.1\%).
 
-The preliminary analyses presented in this project, provide an initial exploration into the uses for these data.
-In future studies, these data could be used to see how the trails change due to human disturbance or presence, or to see how and where elephants do or do not change course.
-There are also opportunities with these data to look at the change of vegetation along the different use trails and if it will affect the reproduction, seed dispersal or the types of plants that come up in the forest gaps.
+The preliminary analyses presented in this project, provide an initial
+exploration into the uses for these data. In future studies, these data
+could be used to see how the trails change due to human disturbance or
+presence, or to see how and where elephants do or do not change course.
+There are also opportunities with these data to look at the change of
+vegetation along the different use trails and if it will affect the
+reproduction, seed dispersal or the types of plants that come up in the
+forest gaps.
+
+\end{document}

### 2021-04-27 draft_writing.Rmd
updates to draft writing doc and ds_store file

@@
 
-#change to data 
+#change to date 
 tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
@@
 ### SUMMARY AND CONCLUSIONS
+The first question we wanted to answer was to see the number of trail points occurring for each day.
+In general, you can see that the level of use increased from poorly trodden to medium trodden over time.
+By looking at the graph, the poor trodden trails are more abundant earlier in the year and the medium trodden trails become more abundant in April and May.
+
 In our final dataset, which included only the data that were recorded on common dates between our two initial datasets, there are 27,218 poor trodden trail points, 13,666 medium trodden trail points, and 26 heavily trodden trail points.
